Typed Functional Genetic Programming

### 2013-07-19 tex/thesis.tex
[tex]

\newcommand{\tur}[3]{#1 \vdash #2 : #3}

\newcommand{\turst}[3]{$#1 \vdash #2 : #3$\xspace}
\newcommand{\GMS}{\turst{\Gamma}{M}{\sigma}}

@@

Notable highlights are: 
\begin{itemize}
 \item \textit{Inhabitation tree} (a slight modification 
   of \textit{inhabitation machine} presented in \cite{barendregt10}
   suited better for reasoning about implementation details) 
   is presented and used for \textit{generation of individuals}.
   This generation process is parameterized by simple 
   \textit{search strategy}. Several search strategies are presented.   
 \item \textit{Abstraction elimination} is utilized
   to enable use of simple \textit{mutation operator}.
 \item \textit{Polymorphic} generalization of inhabitation tree,
   enabling use of polymorphic functions as building blocks, 
   is presented.

@@

@@

\begin{definition}
A statement $M : \sigma$ is \textit{derivable from}
a context $\Gamma$ if it can be produced by the following rules.
\begin{align*}
(x,\sigma) \in \Gamma &~\then~ \tur{\Gamma}{x}{\sigma}\\
\tur{\Gamma}{M}{(\sigma \ar \tau)}~,~\tur{\Gamma}{N}{\sigma} 
&~\then~ \tur{\Gamma}{(M~N)}{\tau}\\  
\tur{\Gamma \cup (x,\sigma)}{M}{\tau}
&~\then~ \tur{\Gamma}{\lamb{x}{M}}{(\sigma \ar \tau)} 
\end{align*}~
\end{definition}

By writing \GMS we say 
{\it statement $M : \sigma$ is derivable from context $\Gamma$ }.

We construct valid statements of the form \GMS 
by using inference rules.\\

	Basically speaking, inference rules are used for deriving statements of a form 
	\GMS from yet derived statements of such a form.
	Those inference rules are written in the following form:

	\begin{equation*}
		\frac{\Gamma_1 \vdash M_1 : \sigma_1 \qquad
			  \Gamma_2 \vdash M_2 : \sigma_2 \quad
			  \dotsm \quad
		      \Gamma_n \vdash M_n : \sigma_n}
		     {\Gamma_{n+1} \vdash M_{n+1} : \sigma_{n+1}}
	\end{equation*}

	Suppose we have yet derived statements 
	$\Gamma_1 \vdash M_1 : \sigma_1 ,
	 \Gamma_2 \vdash M_2 : \sigma_2 ,
	 \dots ,
	 \Gamma_n \vdash M_n : \sigma_n$. 
	It allows as to use the inference rule to derive statement
	\mbox{ $\Gamma_{n+1} \vdash M_{n+1} : \sigma_{n+1}$ }.

	For deriving statements including types of a form 
	$(\sigma \rightarrow \tau)$ are essential those two 
	inference rules:

	\begin{equation*}
		\frac{\Gamma \vdash M : \sigma \rightarrow \tau \qquad
			  \Gamma \vdash N : \sigma }
		     {\Gamma \vdash (M N) : \tau }
	\end{equation*}

	\begin{equation*}
		\frac{\Gamma \cup \{ ( x,\sigma ) \} \vdash M : \tau }
		     {\Gamma \vdash (\lambda x . M) : \sigma \rightarrow \tau }
	\end{equation*}

	This kind of inference rules allows us to derive new statements from yet derived statements, but 
	what if we do not have any statement yet? 
	For this purpose we have other kinds of inference rules such as {\it axiom} inference rule:

	\begin{equation*}
		\frac{( x , \sigma )  \in \Gamma}
		     {\Gamma \vdash x : \sigma}
	\end{equation*}

	\Lets consider an example statement of a form \GMS :

	\[
		\{\} \vdash (\lambda f . (\lambda x . (f x) )) : 
		(\sigma \rightarrow \tau) \rightarrow ( \sigma \rightarrow \tau ) 
	\]

	This statement is derived as follows:

@@

\subsection{Simply typed $\lambda$-calculus}

\begin{todo}
   & tady to nějak okomentovat že to je teda vono
   & říct že se vždycky zastavěj (strong normalization) ale
     asi to říct na jinym místě než tady, tady se teprv definujou typy, 
     říct to někde kde už se ví jak sou napojený na termy    
\end{todo}

\section{Term generating grammar}

Inference rules are good for deriving statements of a form \GMS, but our
goal is slightly different; we would like to generate many \lterms M for a given type 
$\sigma$ and context $\Gamma$.

Our approach will be to take each inference rule and transform it to a rule of term generating
grammar. With this term generating grammar it will be much easier to reason about generating 
\lterms.

It won't be a grammar in classical sense because we will be operating with infinite sets of
nonterminal symbols and rules. \footnote{TODO : mention terminal symbols - situation around 
variables and their construction with ' symbol.}

Let $Non = Type \times Context $ be our {\it nonterminal} set. 
So for every $i \in Non$ is $i = (\sigma_i , \Gamma_i )$.

\Lets consider each relevant inference rule and its corresponding grammar rule.

First inference rule is {\it implication elimination} also known as 
{\it modus ponens}: 
\[
	\frac{\Gamma \vdash M : \sigma \rightarrow \tau \qquad
		  \Gamma \vdash N : \sigma }
	     {\Gamma \vdash (M N) : \tau }
\]
\\
For every $\sigma, \tau \in \mathbb{T}$ and for every {\it context} 
$\Gamma \in \mathfrak P \left({\Lambda \times  \mathbb{T}}\right)$ there is a grammar rule of a form\footnote{ 
Terminal symbols for parenthesis and normally {\it space} now \textvisiblespace \quad (for {\it function application} operator) are visually highlighted. }: 
\[	
	( \tau , \Gamma )  \longmapsto
	\bigg( ( \sigma \rightarrow \tau , \Gamma ) 
	  \mbox{ \Vtextvisiblespace[1em] } ( \sigma , \Gamma ) \bigg)
\]
\\

Second inference rule is {\it implication introduction}: 
\[
	\frac{\Gamma \cup \{ ( x,\sigma ) \} \vdash M : \tau }
	     {\Gamma \vdash (\lambda x . M) : \sigma \rightarrow \tau }
\]
\\
$\forall \sigma, \tau \in \mathbb{T}$ 
$\forall${\it context} $\Gamma \in \mathfrak P \left({\Lambda \times  \mathbb{T}}\right) $ 
$\forall x \in V $ such that there is no $(x,\rho) \in \Gamma$ 
there is a grammar rule:
\[ 
	( \sigma \rightarrow \tau , \Gamma )  \longmapsto
	\bigg( \mbox{ {\Large $\lambda$ x . }}( \tau , \Gamma \cup \{ (x,\sigma) \} ) \quad \bigg)
\]
\\

Third inference rule is {\it axiom}: 
\[
		\frac{( x , \sigma )  \in \Gamma}
		     {\Gamma \vdash x : \sigma}
\]
\\
$\forall \sigma \in \mathbb{T}$ 
$\forall${\it context} $\Gamma \in \mathfrak P \left({\Lambda \times  \mathbb{T}}\right) $ 
$\forall x \in V \cup C $ such that $(x,\sigma) \in \Gamma$ 
there is a grammar rule:
\[ 
	( \sigma , \Gamma )  \longmapsto \mbox{ {\Large x}}
\]
\\

We will demonstrate \lterm generation on example. 
Again on $(\lambda f . (\lambda x . (f x) ))$. 
We would like to generate \lterm of a type 
$(\sigma \rightarrow \tau) \rightarrow (\sigma \rightarrow \tau)$
with $\Gamma = \{\}$.
\begin{align*}
	& ((\sigma \rightarrow \tau) \rightarrow (\sigma \rightarrow \tau),\{\}) \\ 
	\longmapsto & \Big( \mbox{ {\Large $\lambda$f.}}
	  ( \sigma \rightarrow \tau , \{ (f,\sigma \rightarrow \tau) \} ) 
	~ \Big)
	\\
	\longmapsto & 
	\Big( \mbox{ {\Large $\lambda$f. }}
		\Big( \mbox{ {\Large $\lambda$x. }}
	  	 	( \tau , \{ (f,\sigma \rightarrow \tau) , (x,\sigma) \} ) 
		~ \Big)  	 
	~ \Big)
	\\
	\longmapsto & 
	\Big( \mbox{ {\Large $\lambda$f. }}
		\Big( \mbox{ {\Large $\lambda$x. }}	  	 	
	  	 	\Big( 
	  	 	  ( \sigma \rightarrow \tau , \{ (f,\sigma \rightarrow \tau) , (x,\sigma) \} ) 
			  \mbox{ \Vtextvisiblespace[1em] } 
			  ( \sigma , \{ (f,\sigma \rightarrow \tau) , (x,\sigma) \} )  \Big) 
		~ \Big)  	 
	 ~ \Big)
	\\
	\longmapsto & 
	\Big( \mbox{ {\Large $\lambda$f. }}
		\Big( \mbox{ {\Large $\lambda$x. }}	  	 	
	  	 	\Big( 
	  	 	  \mbox{ {\Large f}} 
			  \mbox{ \Vtextvisiblespace[1em] } 
			  ( \sigma , \{ (f,\sigma \rightarrow \tau) , (x,\sigma) \} ) \Big) 
		~ \Big)  	 
	~ \Big)		
	\\
	\longmapsto & 
	\Big( \mbox{ {\Large $\lambda$f. }}
		\Big( \mbox{ {\Large $\lambda$x. }}	  	 	
	  	 	\Big( 
	  	 	  \mbox{ {\Large f}} 
			  \mbox{ \Vtextvisiblespace[1em] } 
			  \mbox{{\Large x}} \Big) 
		~ \Big)  	 
	~ \Big)
\end{align*}

@@

Let \GMS where $\sigma = \tau_0 \ar \tau_1 \ar \dots \ar \tau_n \ar \alpha$.\\
Then $M$ is in \textit{long normal form} if following 
conditions are satisfied:

>>>>

\newcommand{\tur}[3]{#1\vdash{}#2:#3}

\newcommand{\turst}[3]{$#1\vdash{}#2:#3$\xspace}
\newcommand{\GMS}{\turst{\Gamma}{M}{\sigma}}

@@

Notable highlights are: 
\begin{itemize}
 \item \textit{Inhabitation tree} (a slight modification 
   of \textit{inhabitation machine} presented in \cite{barendregt10}
   suited better for reasoning about implementation details) 
   is presented and used for \textit{generation of individuals}.
   This generation process is parameterized by simple 
   \textit{search strategy}. Several search strategies are presented,
   such as \textit{systematic} strategy and strategy performing
   classical GP individual generation.     
 \item \textit{Abstraction elimination} is utilized
   to enable use of simple \textit{mutation operator}.
 \item \textit{Polymorphic} generalization of inhabitation tree,
   enabling use of polymorphic functions as building blocks, 
   is presented.

@@

\begin{todo}
& zmínit $=_\beta$ a $\twoheadrightarrow_\beta$ 
jestli se k tomu někde dál v textu odkazuju
\end{todo}

@@

\begin{definition}
A statement $M\colon\sigma$ is \textit{derivable from}
a context $\Gamma$ (notation $\Gamma\vdash{}M\colon\sigma$) 
if it can be produced by the following rules.
\begin{align*}
(x,\sigma) \in \Gamma &~\then~ \tur{\Gamma}{x}{\sigma}\\
\tur{\Gamma}{M}{\sigma \ar \tau}~,~\tur{\Gamma}{N}{\sigma} 
&~\then~ \tur{\Gamma}{(M~N)}{\tau}\\  
\tur{\Gamma \cup \{(x,\sigma)\}}{M}{\tau}
&~\then~ \tur{\Gamma}{\lamb{x}{M}}{\sigma \ar \tau} 
\end{align*}~
\end{definition}

Those rules are usually stated in the form of \textit{inference rules}.

\newcommand{\ruleI}[2]{\dfrac{#1}{#2}}
\newcommand{\ruleII}[3]{\dfrac{#1 \qquad #2}{#3}}
\newcommand{\ruleIII}[4]{\dfrac{#1 \qquad #2 \qquad #3}{#4}}
\newcommand{\ruleIV}[5]{\dfrac{#1 \qquad #2 \qquad #3 \qquad #4}{#5}}

\begin{align*}
\ruleI { (x:\sigma) \in \Gamma }
       { \tur{\Gamma}{x}{\sigma} }
&& \textit{(axiom)} \\[1em]
\ruleII{ \tur{\Gamma}{M}{\sigma \ar \tau} }
       { \tur{\Gamma}{N}{\sigma} }
       { \tur{\Gamma}{(M~N)}{\tau} }
&& \textit{($\ar$-elimination)} \\[1em]
\ruleI { \tur{\Gamma,x:\sigma}{M}{\tau} }
       { \tur{\Gamma}{\lamb{x}{M}}{\sigma \ar \tau} }
&& \textit{($\ar$-introduction)} \\[1em]
\end{align*}

\Lets consider following example statement of a form \GMS.

$$
	\{\} \vdash (\lambda f . (\lambda x . (f x) )) : 
	(\sigma \rightarrow \tau) \rightarrow ( \sigma \rightarrow \tau ) 
$$~

This statement is derived as follows.

@@

If we turn the derivation upside down, we can see that it is a tree with 
desired statement in the root and \textit{axioms} in the leafs. \\

New valid inference rules may be produced by finite composition of 
other valid inference rules.

\Lets consider following rule.

$$
\ruleII
 { \tur{\Gamma,x:\sigma}{M}{\tau} }
 { \tur{\Gamma}{N}{\sigma} }
 { \tur{\Gamma}{(\lamb{x}{M}~N)}{\tau} }
$$

Its validity is shown by the following rule composition.

$$
\ruleII
	{ \ruleI 
	     { \tur{\Gamma,x:\sigma}{M}{\tau} }
	     { \tur{\Gamma}{\lamb{x}{M}}{\sigma \ar \tau} }
	}
	{ \tur{\Gamma}{N}{\sigma} }
	{ \tur{\Gamma}{(\lamb{x}{M}~N)}{\tau} }
$$~

\Lets consider following strange rule.

$$
\ruleII
 { \tur{\Gamma,x:\rho \ar \pi}{M}{\tau} }
 { \tur{\Gamma,y:\rho}{N}{\pi} }
 { \tur{\Gamma}{(\lamb{x}{M}~\lamb{y}{N})}{\tau} }
$$

Its validity is shown by the following rule composition.

$$
\ruleII
	{ \ruleI 
	     { \tur{\Gamma,x:\rho \ar \pi}{M}{\tau} }
	     { \tur{\Gamma}{\lamb{x}{M}}{(\rho \ar \pi) \ar \tau} }
	}
	{ \ruleI 
	     { \tur{\Gamma,y:\rho}{N}{\pi} }
	     { \tur{\Gamma}{\lamb{y}{N}}{\rho \ar \pi} }
	}	 { \tur{\Gamma}{(\lamb{x}{M}~\lamb{y}{N})}{\tau} }
$$~

Generally speaking, inference rules are used for deriving statements of a form 
\GMS from yet derived statements of such a form (with exception of 
\textit{axiom} rule, which serves as a "starting point").
Those inference rules are written in the following form.

$$ 
\ruleIV
  { \tur{\Gamma_1}{M_1}{\sigma_1} }
  { \tur{\Gamma_2}{M_2}{\sigma_2} }
  { \dotsm }
  { \tur{\Gamma_n}{M_n}{\sigma_n} }
  { \tur{\Gamma_{n+1}}{ \mathbf{T}(M_1,M_2,\dots,M_n)}{\sigma_{n+1}} } 
$$

Where $\mathbf{T}(M_1,M_2,\dots,M_n)$ stands for some 
specific \lterm that may contain \textit{"meta-variables"}
$M_1,M_2,\dots,M_n$.

In the example above $\mathbf{T}(M,N) = (\lamb{x}{M}~\lamb{y}{N})$

Suppose we have derived statements 
$\Gamma_1 \vdash M_1 : \sigma_1 ,
 \Gamma_2 \vdash M_2 : \sigma_2 ,
 \dots ,
 \Gamma_n \vdash M_n : \sigma_n$. 
Then we are allowed to use the inference rule to derive statement
\mbox{ $\Gamma_{n+1} \vdash \mathbf{T}(M_1,M_2,\dots,M_n) : \sigma_{n+1}$ }.\\

%For deriving statements including types of a form 
%$(\sigma \rightarrow \tau)$ are essential those two 
%inference rules:
%
%\begin{equation*}
%	\frac{\Gamma \vdash M : \sigma \rightarrow \tau \qquad
%		  \Gamma \vdash N : \sigma }
%	     {\Gamma \vdash (M N) : \tau }
%\end{equation*}	
%
%\begin{equation*}
%	\frac{\Gamma \cup \{ ( x,\sigma ) \} \vdash M : \tau }
%	     {\Gamma \vdash (\lambda x . M) : \sigma \rightarrow \tau }
%\end{equation*}		 
% 
%This kind of inference rules allows us to derive new statements from yet derived statements, but 
%what if we do not have any statement yet? 
%For this purpose we have other kinds of inference rules such as {\it axiom} inference rule:   
%	
%\begin{equation*}
%	\frac{( x , \sigma )  \in \Gamma}
%	     {\Gamma \vdash x : \sigma}
%\end{equation*}

\begin{todo}
   & říct že se vždycky zastavěj (strong normalization),
     && potřebný definice pro to uvést v \lterms sekci
   & možná zmínit že se jedná o \textit{gentzen style proof}
   & říct že jsou to modus ponens a jeden směr věty o dedukci  
\end{todo}

\section{Term generating grammar}
Inference rules are good for deriving statements of the form \GMS, but our
goal is slightly different; we would like to generate many \lterms M for a given type 
$\sigma$ and context $\Gamma$.\\

\Lets start our reasoning about generation of \lterms with
notion of term generating grammar. We will show how to
transform an inference rule into a rule of term generating grammar.
With the appropriate term generating grammar 
in hands it is just a simple step to \textit{inhabitation trees}.\\

It is not a grammar in classical sense, because we will be operating with infinite sets of nonterminal symbols and rules. Technically it it
2-level grammer \cite{todo}.\\

Our terminals are symbols used for writing \lterms 
.%,i.e., parentheses,  .

Our non-terminals are pairs $(\tau,\Gamma)$ where 
$\tau \in \T$ and $\Gamma$ is a \textit{context}.

Our initial non-terminal is $S = (\sigma_S,\Gamma_S)$
where $\sigma_S$ is desired type of generated \lterms and
$\Gamma_S$ is \textit{building blocks} context. \\

\Lets begin with \textit{axiom} inference rule.

$$
\ruleI { (x:\sigma) \in \Gamma }
       { \tur{\Gamma}{x}{\sigma} }
$$~

%It corresponds to following grammar rule schema.
%$$ ( \sigma , \Gamma )  \longmapsto \mbox{ {\Large x}} 
%~~~~~~~~~~~~~ \textbf{if } (x,\sigma) \in \Gamma$$

For every non-terminal $(\sigma,\Gamma)$ 
and for every $x$ such that $(x:\sigma) \in \Gamma$,
there is following grammar rule.
$$ ( \sigma , \Gamma )  \longmapsto \mbox{ {\Large x}} $$ \\

Next is \textit{$\ar$-elimination}.

$$
\ruleII{ \tur{\Gamma}{M}{\sigma \ar \tau} }
       { \tur{\Gamma}{N}{\sigma} }
       { \tur{\Gamma}{(M~N)}{\tau} }
$$~

For every non-terminal $(\tau,\Gamma)$ 
and for every type $\sigma \in \T$ 
there is following grammar rule.\footnote{ 
Terminal symbols for parenthesis and normally {\it space} 
now \textvisiblespace \quad (for {\it function application} operator) 
are visually highlighted. } 
$$
	( \tau , \Gamma )  \longmapsto
	\bigg( ( \sigma \rightarrow \tau , \Gamma ) 
	\mbox{ \Vtextvisiblespace[1em] } ( \sigma , \Gamma ) \bigg)
$$\\

And finally {\it $\ar$-introduction}. 
$$
\ruleI { \tur{\Gamma,x:\sigma}{M}{\tau} }
       { \tur{\Gamma}{\lamb{x}{M}}{\sigma \ar \tau} }
$$

For every non-terminal $(\sigma \ar \tau,\Gamma)$
and for every variable $x$
there is following grammar rule.\footnote{
\textbf{TODO !!!!} nějak okomentovat že by měla být freš,
pokudmožno v celym termu ale to že byse dělalo blě aby to
zůstala bezkontextová gramatika, ale je to celkem nepodtatnej detail
protože tyhle term generating gramatiky sou tu spíš
pro zajímavost a pro pevnější kramfleky, než pro praktický 
generování}

%and let $x$ be variable name that is
%not present in any declaration in $\Gamma$.
%(since we want to generate "fresh" variables).

$$ 
	( \sigma \rightarrow \tau , \Gamma )  \longmapsto
	\bigg( \mbox{ {\Large $\lambda$ x . }}( \tau , \Gamma \cup \{ (x,\sigma) \} ) \quad \bigg)
$$\\

We will demonstrate \lterm generation on example. 
Again on $(\lambda f . (\lambda x . (f x) ))$. 
We would like to generate \lterm of a type 
$(\sigma \rightarrow \tau) \rightarrow (\sigma \rightarrow \tau)$
with $\Gamma = \{\}$.
\begin{align*}
	& ((\sigma \rightarrow \tau) \rightarrow (\sigma \rightarrow \tau),\{\}) \\ 
	\longmapsto & \Big( \mbox{ {\Large $\lambda$f.}}
	  ( \sigma \rightarrow \tau , \{ (f,\sigma \rightarrow \tau) \} ) 
	~ \Big)
	\\
	\longmapsto & 
	\Big( \mbox{ {\Large $\lambda$f. }}
		\Big( \mbox{ {\Large $\lambda$x. }}
	  	 	( \tau , \{ (f,\sigma \rightarrow \tau) , (x,\sigma) \} ) 
		~ \Big)  	 
	~ \Big)
	\\
	\longmapsto & 
	\Big( \mbox{ {\Large $\lambda$f. }}
		\Big( \mbox{ {\Large $\lambda$x. }}	  	 	
	  	 	\Big( 
	  	 	  ( \sigma \rightarrow \tau , \{ (f,\sigma \rightarrow \tau) , (x,\sigma) \} ) 
			  \mbox{ \Vtextvisiblespace[1em] } 
			  ( \sigma , \{ (f,\sigma \rightarrow \tau) , (x,\sigma) \} )  \Big) 
		~ \Big)  	 
	 ~ \Big)
	\\
	\longmapsto & 
	\Big( \mbox{ {\Large $\lambda$f. }}
		\Big( \mbox{ {\Large $\lambda$x. }}	  	 	
	  	 	\Big( 
	  	 	  \mbox{ {\Large f}} 
			  \mbox{ \Vtextvisiblespace[1em] } 
			  ( \sigma , \{ (f,\sigma \rightarrow \tau) , (x,\sigma) \} ) \Big) 
		~ \Big)  	 
	~ \Big)		
	\\
	\longmapsto & 
	\Big( \mbox{ {\Large $\lambda$f. }}
		\Big( \mbox{ {\Large $\lambda$x. }}	  	 	
	  	 	\Big( 
	  	 	  \mbox{ {\Large f}} 
			  \mbox{ \Vtextvisiblespace[1em] } 
			  \mbox{{\Large x}} \Big) 
		~ \Big)  	 
	~ \Big)
\end{align*}~\\

\Lets consider a general inference rule again.
$$ 
\ruleIV
  { \tur{\Gamma_1}{M_1}{\sigma_1} }
  { \tur{\Gamma_2}{M_2}{\sigma_2} }
  { \dotsm }
  { \tur{\Gamma_n}{M_n}{\sigma_n} }
  { \tur{\Gamma_{n+1}}{ \mathbf{T}(M_1,M_2,\dots,M_n)}{\sigma_{n+1}} } 
$$

The corresponding grammar rule can be summarized as follows.

$$
 (\sigma_{n+1},\Gamma_{n+1})
 \longmapsto
 \mbox{ {\Large \textbf{T}}} 
 \bigg( 
 (\sigma_1,\Gamma_1)
 \mbox{ {\Large ,}}
 (\sigma_2,\Gamma_2)
 \mbox{ {\Large ,}}
 ~\dots
 \mbox{ {\Large ,}}
 (\sigma_n,\Gamma_n)
 \bigg)  
$$

\begin{todo}
   & asi ocitovat toho van Wijngaarden [1981] s těma 2-level grammar
     jak je to v Barendregtovi \cite{barendregt10} 
   & vyřešit jak je to stou freš proměnnou x v introdukci...
     && stejně tak v tom obecnym případě to dělá neplechu
\end{todo}

%\section{Term generating grammar}
%
%Inference rules are good for deriving statements of the form \GMS, but our
%goal is slightly different; we would like to generate many \lterms M for a given type 
%$\sigma$ and context $\Gamma$.
%
%Our approach will be to take each inference rule and transform it to a rule of term generating
%grammar. With this term generating grammar it will be much easier to reason about generating 
%\lterms.
%	
%It won't be a grammar in classical sense because we will be operating with infinite sets of
%nonterminal symbols and rules. \footnote{TODO : mention terminal symbols - situation around 
%variables and their construction with ' symbol.}
%
%Let $Non = Type \times Context $ be our {\it nonterminal} set. 
%So for every $i \in Non$ is $i = (\sigma_i , \Gamma_i )$.
%
%\Lets consider each relevant inference rule and its corresponding grammar rule.
%
%First inference rule is {\it implication elimination} also known as 
%{\it modus ponens}: 
%\[
%	\frac{\Gamma \vdash M : \sigma \rightarrow \tau \qquad
%		  \Gamma \vdash N : \sigma }
%	     {\Gamma \vdash (M N) : \tau }
%\]
%\\
%For every $\sigma, \tau \in \mathbb{T}$ and for every {\it context} 
%$\Gamma \in \mathfrak P \left({\Lambda \times  \mathbb{T}}\right)$ there is a grammar rule of a form\footnote{ 
%Terminal symbols for parenthesis and normally {\it space} now \textvisiblespace \quad (for {\it function application} operator) are visually highlighted. }: 
%\[	
%	( \tau , \Gamma )  \longmapsto
%	\bigg( ( \sigma \rightarrow \tau , \Gamma ) 
%	  \mbox{ \Vtextvisiblespace[1em] } ( \sigma , \Gamma ) \bigg)
%\]
%\\
%
%Second inference rule is {\it implication introduction}: 
%\[
%	\frac{\Gamma \cup \{ ( x,\sigma ) \} \vdash M : \tau }
%	     {\Gamma \vdash (\lambda x . M) : \sigma \rightarrow \tau }
%\]
%\\
%$\forall \sigma, \tau \in \mathbb{T}$ 
%$\forall${\it context} $\Gamma \in \mathfrak P \left({\Lambda \times  \mathbb{T}}\right) $ 
%$\forall x \in V $ such that there is no $(x,\rho) \in \Gamma$ 
%there is a grammar rule:
%\[ 
%	( \sigma \rightarrow \tau , \Gamma )  \longmapsto
%	\bigg( \mbox{ {\Large $\lambda$ x . }}( \tau , \Gamma \cup \{ (x,\sigma) \} ) \quad \bigg)
%\]
%\\	
%
%Third inference rule is {\it axiom}: 
%\[
%		\frac{( x , \sigma )  \in \Gamma}
%		     {\Gamma \vdash x : \sigma}
%\]
%\\
%$\forall \sigma \in \mathbb{T}$ 
%$\forall${\it context} $\Gamma \in \mathfrak P \left({\Lambda \times  \mathbb{T}}\right) $ 
%$\forall x \in V \cup C $ such that $(x,\sigma) \in \Gamma$ 
%there is a grammar rule:
%\[ 
%	( \sigma , \Gamma )  \longmapsto \mbox{ {\Large x}}
%\]
%\\
%
%We will demonstrate \lterm generation on example. 
%Again on $(\lambda f . (\lambda x . (f x) ))$. 
%We would like to generate \lterm of a type 
%$(\sigma \rightarrow \tau) \rightarrow (\sigma \rightarrow \tau)$
%with $\Gamma = \{\}$.
%\begin{align*}
%	& ((\sigma \rightarrow \tau) \rightarrow (\sigma \rightarrow \tau),\{\}) \\ 
%	\longmapsto & \Big( \mbox{ {\Large $\lambda$f.}}
%	  ( \sigma \rightarrow \tau , \{ (f,\sigma \rightarrow \tau) \} ) 
%	~ \Big)
%	\\
%	\longmapsto & 
%	\Big( \mbox{ {\Large $\lambda$f. }}
%		\Big( \mbox{ {\Large $\lambda$x. }}
%	  	 	( \tau , \{ (f,\sigma \rightarrow \tau) , (x,\sigma) \} ) 
%		~ \Big)  	 
%	~ \Big)
%	\\
%	\longmapsto & 
%	\Big( \mbox{ {\Large $\lambda$f. }}
%		\Big( \mbox{ {\Large $\lambda$x. }}	  	 	
%	  	 	\Big( 
%	  	 	  ( \sigma \rightarrow \tau , \{ (f,\sigma \rightarrow \tau) , (x,\sigma) \} ) 
%			  \mbox{ \Vtextvisiblespace[1em] } 
%			  ( \sigma , \{ (f,\sigma \rightarrow \tau) , (x,\sigma) \} )  \Big) 
%		~ \Big)  	 
%	 ~ \Big)
%	\\
%	\longmapsto & 
%	\Big( \mbox{ {\Large $\lambda$f. }}
%		\Big( \mbox{ {\Large $\lambda$x. }}	  	 	
%	  	 	\Big( 
%	  	 	  \mbox{ {\Large f}} 
%			  \mbox{ \Vtextvisiblespace[1em] } 
%			  ( \sigma , \{ (f,\sigma \rightarrow \tau) , (x,\sigma) \} ) \Big) 
%		~ \Big)  	 
%	~ \Big)		
%	\\
%	\longmapsto & 
%	\Big( \mbox{ {\Large $\lambda$f. }}
%		\Big( \mbox{ {\Large $\lambda$x. }}	  	 	
%	  	 	\Big( 
%	  	 	  \mbox{ {\Large f}} 
%			  \mbox{ \Vtextvisiblespace[1em] } 
%			  \mbox{{\Large x}} \Big) 
%		~ \Big)  	 
%	~ \Big)
%\end{align*}
%

@@

Let \GMS where 
$\sigma = \tau_0 \ar \tau_1 \ar \dots \ar \tau_n \ar \alpha$,
$n \geq 0$.\\
Then $M$ is in \textit{long normal form} if following 
conditions are satisfied:
